Language Skills in an Ethnically Segmented Labour market: Estonia 1989 -- 2012

### 2018-07-22 paper/ijm.tex
abstract rewritten

%\maketitle
%\begin{center}
%	\textbf{{\large Language Skills in an Ethnically Segmented Labour market:
%				Estonia 1989 -- 2012}} \\
%	{\large Sven-Kristjan Bormann} \\
%	\emph{School of Economics and Business Administration, University of Tartu, Narva mnt. 4, 51007 Tartu, Estonia\\ Tel.: +372 5844 3873\\ Email: \href{mailto:bormann@ut.ee}{bormann@ut.ee}}	\\
%
%	{\large Svetlana Ridala} \\
%	\emph{School of Business and Governance, Tallinn University of Technology,  Akadeemia tee 3, 12618 Tallinn, Estonia\\ Email: \href{
%			mailto:svetlana.ridala@gmail.com}{
%			svetlana.ridala@gmail.com}} \\
%		{\large Ott Toomet}\\
%		\emph{Information School, University of Washington, USA; Email: \href{mailto:otoomet@uw.edu}{otoomet@uw.edu}}
%%	\emph{(Received XX XXXX 20XX; accepted XX XXXX 20XX)}\\
%	\textbf{Corresponding author:} Sven-Kristjan Bormann Email: \href{mailto:bormann@ut.ee}{bormann@ut.ee}
%\end{center}
%\paragraph{Short biographical note:} \hspace{0pt} \\
%Sven-Kristjan Bormann born 28.11.1985, graduated in 2011 from Kiel University, Germany in Economics with specialisation in Quantitative Economics and Computer Science. Since September 2012 PhD-student in Economics at Tartu University, Estonia. \\
%Svetlana Ridala born 28.03.1976, graduated in 2003 from Tallinn Pedagogical University, Estonia master's (scien) degree in mathematics. Since 2009 PhD-student in Economics at Tallinn School of Economics and Business Administration, Estonia.
%\\
%Ott-Siim Toomet born 18.03.1970, graduated in 2004 from Aarhus University, Denmark PhD in Economics. Since 2015 Visiting scholar, Information School, University of Washington.
%
%\paragraph{Research funding}
%This research was supported by European Social Fund's Doctoral Studies
%and Internationalisation Program DoRa through Archimedes Foundation, and the Estonian Science Foundation grant 8627.
%\newpage
%\noindent
{\centering \textbf{Language Skills in an Ethnically Segmented Labour market: Estonia 1989 -- 2012}}
\begin{abstract}

	\noindent	\textbf{Purpose --} The purpose of this paper is to study the relationship between language skills and labour market
outcomes in Estonia, a linguistically divided country where Estonian
is the sole official language.\\
\textbf{Design/methodology/approach --} The authors use the Estonian Labour Force Survey from 1989 to 2012 to study the impact of language skills on unemployment (probability) and wages.\\
\textbf{Findings --} For men, fluency in                                   
Estonian is related to approximately 5 percentage points lower unemployment, but to virtually no income
premium. English skills are associated with a
substantial income premium, but are virtually unrelated to
unemployment. Russian skills are related to higher wages for Estonians,
but the premium is falling over time and Russian is not associated with lower unemployment.	\\
\textbf{Research limitations/implications -- } No casual claims are possible unless a suitable instruments for language skills can be found. Individual characteristics are able to explain a
large part of English returns, so that part of the remaining effect
may also be related to an ability bias. 
The usage of self-reported language skills may introduce measurement errors, which may	decrease in the future, as language test data becomes available.\\	
\textbf{Practical implications --} These results can be explained by simultaneous ethnic and gender
segregation where the high-skilled male labour market is more segregated
than the female labour market, and females are more often working in
communication-intensive jobs.\\
\textbf{Originality/value --} This paper provides a comprehensive overview about the effects of language skills on labour market outcomes for a linguistically divided, low immigration population. In contrast, the existing literature deals mostly with the effects of language skills on the labour market outcomes of immigrants. \\
\end{abstract}
\textbf{Keywords:} language skills, ethnicity, wage gap, unemployment, Estonia \\
\textbf{JEL classification:} J15, J31, J71

>>>>

%\maketitle
%\begin{center}
%	\textbf{{\large Language Skills in an Ethnically Segmented Labour market:
%				Estonia 1989 -- 2012}} \\
%	{\large Sven-Kristjan Bormann} \\
%	\emph{School of Economics and Business Administration, University of Tartu, Narva mnt. 4, 51007 Tartu, Estonia\\ Tel.: +372 5844 3873\\ Email: \href{mailto:bormann@ut.ee}{bormann@ut.ee}}	\\
%
%	{\large Svetlana Ridala} \\
%	\emph{School of Business and Governance, Tallinn University of Technology,  Akadeemia tee 3, 12618 Tallinn, Estonia\\ Email: \href{
%			mailto:svetlana.ridala@gmail.com}{
%			svetlana.ridala@gmail.com}} \\
%		{\large Ott Toomet}\\
%		\emph{Information School, University of Washington, USA; Email: \href{mailto:otoomet@uw.edu}{otoomet@uw.edu}}
%%	\emph{(Received XX XXXX 20XX; accepted XX XXXX 20XX)}\\
%	\textbf{Corresponding author:} Sven-Kristjan Bormann Email: \href{mailto:bormann@ut.ee}{bormann@ut.ee}
%\end{center}
%\paragraph{Short biographical note:} \hspace{0pt} \\
%Sven-Kristjan Bormann born 28.11.1985, graduated in 2011 from Kiel University, Germany in Economics with specialisation in Quantitative Economics and Computer Science. Since September 2012 PhD-student in Economics at Tartu University, Estonia. \\
%Svetlana Ridala born 28.03.1976, graduated in 2003 from Tallinn Pedagogical University, Estonia master's (scien) degree in mathematics. Since 2009 PhD-student in Economics at Tallinn School of Economics and Business Administration, Estonia.
%\\
%Ott-Siim Toomet born 18.03.1970, graduated in 2004 from Aarhus University, Denmark PhD in Economics. Since 2015 Visiting scholar, Information School, University of Washington.
%
%\paragraph{Research funding}
%This research was supported by European Social Fund's Doctoral Studies
%and Internationalisation Program DoRa through Archimedes Foundation, and the Estonian Science Foundation grant 8627.
%\newpage
%\noindent
{\centering \textbf{Language Skills in an Ethnically Segmented Labour market: Estonia 1989 -- 2012}}
\begin{abstract}
  \noindent
  \textbf{Purpose --} We analyze the relationship between skills of
  Estonian, Russian and English, and labour market
  outcomes in Estonia, a linguistically divided country.
  \\
  \textbf{Design/methodology/approach --} We use the Estonian Labour
  Force Surveys 1992--2012.  We rely on multivariate linear regression
  models to document the relationship between language skills and
  labor market outcomes.
  \\
  \textbf{Findings --} Estonian language knowledge (for ethnic Russians)
  are important determinants of unemployment.  Wage, in contrary, is
  closely related to English skills.  Ethic Russian men do not earn
  any premium from speaking Estonian, while women, fluent in Estonian
  earn approximately 10\% more.  For ethnic Estonians, Russian fluency
  is associated with a similar income gain.
  \\
  \textbf{Research limitations/implications -- } Due to the
  observational nature of the data, the effects reported in our study
  are not causal effects.  As a second limitation, the self-reported
  language skills data may be imprecise and hence the effects we
  report may be too small.
  \\
  \textbf{Practical implications --} Our results stress the role of
  workplace segregation, both along gender and ethnic lines, in
  determining the individual labor market experience.  
  \\
  \textbf{Originality/value --} We provide a comprehensive overview of
  the effects of language skills in a rapidly developing labor
  market in a linguistically divided economy.  We analyze several
  languages with different legal status and document long-term
  trends in the effects.
  \\
\end{abstract}
\textbf{Keywords:} language skills, ethnicity, wage gap, unemployment, Estonia \\
\textbf{JEL classification:} J15, J31, J71
